### 2016-12-15 paper.tex
Updated sig extr

\documentclass[twocolumn, 11pt]{paper}
\usepackage[xcolor]{changebar} 
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{listings}

\usepackage{tikz}
\usetikzlibrary{automata,positioning,arrows}

%opening
\title{Automatic extraction of symbolic network signatures from malware code}
\author{Arjun Gurumurthy, Kausik Subramanian}

\begin{document}

\maketitle

\begin{abstract}
We propose the automatic extraction of symbolic network signatures from malware code to aid and automate network-based malware detection systems.
\end{abstract}

\section{Motivation}
Malware, short for malicious software, 
is one of the primary threats to modern-day networks, especially
with the mass proliferation of the Internet. These are used to 
disrupt computer operations, extract sensitive information, and 
gain access to private systems.  Infected hosts can participate 
in distributed denial-of-service attacks (DDoS), these class of 
attacks are very hard to detect and mitigate, leading to immense
loss of revenues. A new shift is towards creating malware for 
profit called botnets, in which a customer can buy a network of
infected hosts for a price. In such a setting, the malware is
specifically designed to evade detection, thus motivating the need
for better detection tools for malware.

The most prevalent form of malware detection is endhost-based detection
(commonly known as anti-virus software). Detected malware's code is hashed
and the code signatures are uploaded to a central repository. By maintaining
an up-to-date repository of malware signatures, the endhost software can
detect malware by comparing program signatures with the repository. However, 
endhost-based malware detection can fail due to various reasons~\cite{networksig}:
\begin{itemize}
	\item \textbf{Anti-virus software is not operational or up to date.} 
	Due to this, the endhost machine can be infected by malware without any
	noticeable effects, thus the user will be oblivious of the infection.
	\item \textbf{Packer programs.} To render programs stealthy, malware authors
	employ packer programs~\cite{packer}. Packers change the program content so that
	its signature differs but its functionality has not changed, thus avoiding
	detection by endhost signature-based software.
\end{itemize}

Instead, a intrusion-detection mechanism integrated 
directly into the service provider’s network offers a
much-needed additional layer of protection. While the malware
code is different, the protocol of the malware used to talk to its
control server (which issues orders to the malware) will not vary. 
Thus, by generating accurate \emph{network signature} of the malware (which
can informally defined as the set of messages sent and received by the malware to
initiate an action), we can use the intrusion detection system in the
network (for e.g., Bro~\cite{bro}) to search for these particular signatures,
and can be used to infer the presence of the malware in the network. However,
manually inspecting malware code is an intractable approach (signatures may 
need to be extracted in real-time). Thus, we propose \emph{automatic extraction 
of accurate network signatures from malware code using different static
analysis techniques.}

\section{Challenges}
We abstract the problem to the following: given a program point,
can we automatically provide an ordered set of symbolic packets 
that must be received (and sent) by the client such that the program 
reaches the program point. The symbolic packets can be used as
network signatures and used to detect malware (a set of packets
satisfying the symbolic packets in order will infer that there 
is a malware with high confidence). By constructing symbolic packet
signatures, we can generalise the intrusion policy and reduce 
true negatives (thus not affecting legitimate traffic).

Malware code can be thought of as a program reacting based on input received
packets received (possibly multi-threaded), which
could require changes to existing reachability analyses tools 
like Codesurfer~\cite{codesurfer}. To extract symbolic network
signatures, we need to define the abstract semantics of the program
suited for network signatures, and define the \emph{combine} and 
\emph{extend} operators used for interprocedural analysis.

%\section{Project Proposal}
%\textbf{Tasks}:
%\begin{itemize}
%	\item Identify and characterize features of common malware code - 1-2 weeks
%	\added{Refer to sections A and B}
%	\item \deleted{Develop the abstract semantics and operators for extracting  
%	symbolic network signatures from malware code - 2-3 weeks} \\
%	\added{Automatically generate a symbolic automaton of the given program
%		based on packet and variable state - 2-3 weeks}
%	\item Modifying/Extending existing tools like Codesurfer for symbolic 
%	network signature extraction - 2-3 weeks
%\end{itemize}
%
%\noindent\textbf{Open source tools}: SMT solvers, malware repositories
%
%\noindent\textbf{Evaluation}: 
%The primary metrics of evaluation are time taken for analysis, accuracy and 
%generality of the symbolic signatures. Performance analysis can be performed 
%with different malware programs with varying lines of code to understand
%the time complexity of the analysis. The quality of the symbolic network
%signatures is a tougher metric to measure. Ideally, the symbolic network 
%signature must only include packets which will make the program reach the
%specified point. If the signature is more general, legitimate packets may 
%be considered as malicious and blocked. On the other hand, the signature
%may miss actual malware traffic because the signature does not encompass
%all inputs leading to the point. We would try to address these trade-offs in
%our evaluation.

\section{Malware program model} \label{sec:malwaremodel}
The malware program listens to incoming packets, and the
network signature is the set of incoming packets which triggers
an attack by the program. We propose a reasonable abstraction
of the code of malware programs.

We model a malware program as a infinite reactive system 
of the form:
\begin{lstlisting}	
while(True) {
	packet = socket.receive();
	...
}
\end{lstlisting}
Each iteration of the loop, the program receives a
packet, and based on current state and packet received,
it performs certain actions. Using static analysis
techniques, we can identify two set of variables, \emph{state}
and \emph{packet} variables. State variables represent the 
program state in terms of the protocol, which packet variables
are used to represent the current packet being processed by
the program.
\begin{lstlisting}	
while(True) {
	packet = socket.receive();
	...
	if (a == 1 and packet[8] = "b") {
		a = a + 1;
	}
	...
}
\end{lstlisting}	
In the above code snippet, $a$ is a state variable
and $packet$ is a packet variable. This also
illustrates the packet processing 
function of the program, which based on the current
input packet and state variables, either transitions
to a new program state or performs an action (attack or
send packets). 
\begin{lstlisting}	
a = 0;
while(True) {
packet = socket.receive();
	if (a == 0 and packet[0] == "a") {
		a = a + 1;
	}
	if (a == 1 and packet[0] == "b") {
		a = a + 1;
	}
	if (a == 2 and packet[0] == "c") {
		attack()  [the attack point is provided as 
				input to analysis]
	}	
}
\end{lstlisting}	
The above code snippet represents the complete model
of the malware program, where based on input packets and
current state, the program transitions to a new state. 
%
%The symbolic automaton for this program is 
%shown in \Cref{fig:progautomaton}. Network signatures
%for the malware are all paths from start state ($a=0$)
%to state of attack. In this case, the sequence of packets
%in order to initiate an attack is $a* \rightarrow b* \rightarrow c*$.
%\begin{figure}[h]
%	\centering
%	\begin{tikzpicture}[shorten >=0.5pt,node distance=3cm,on grid,auto] 
%	\node[state, initial] (v0)   {$a = 0$}; 
%	\node[state] (v1) [right=of v0] {$a = 1$}; 
%	\node[state] (v2) [right=of v1] {$a = 2$};
%	\node[state] (v3) [right=of v2] {$attack$}; 	
%	\path[->] 
%	(v0) edge node {$a*$} (v1)
%	(v1) edge node {$b*$} (v2)
%	(v2) edge node {$c*$} (v3);
%	\end{tikzpicture}
%	\caption{Symbolic automaton for malware program.}
%	\label{fig:progautomaton}
%\end{figure}

\section{Related Work}
An important assumption that we make is that of the simplification of the botnet client code to a set of \textbf{if (state, packet)} statements. Looking at existing research, significant work has gone into automatic protocol reverse-engineering of botnet C\&C protocols using dynamic binary analysis. This generally follows a two step process: Extract the protocol grammar that captures the structure of messages comprising the protocol. Knowing the protocol message format would enable the construction of the protocol state machine, which represents the sequence of messages as a specification of protocol states and transition between states (based on messages received).

Polyglot~\cite{polyglot} is a system that automatically extracts the protocol message format using dynamic binary analysis. The intuition behind the design of the system is that the  way  that  an  implementation  of  the protocol processes the received application data  reveals a wealth of information about the protocol.

Dispatcher~\cite{dispatcher} attempts to perform protocol reverse-engineering for active botnet infiltration. Essentially, it is a tool that makes use of Polyglot to extract the message format and in addition infers the field semantics for messages sent and received by the application. Thus, having knowledge of the network packets received would enable us to make inferences on how they could transition to an 'attack point'.

Prospex~\cite{prospex} is a system for protocol specification extraction for stateful network protocols - essentially, it reverse engineers the protocol state machine by clustering similar network messages and tracks protocol state changes. By merging similar states, a minimal state machine is produced.

All three systems use dynamic analysis techniques - mainly taint propagation. However, this would mean the code would need to execute in a sandboxed environment or be a part of a honeypot system. Static analysis techniques would be a better approach.

Our work focuses on static analysis of the malware code, without having to execute it or capture network traces.

\section{Live variable analysis} \label{sec:livevariableanalysis}
We obtain network signatures using packet and state variables, the complexity increasing with respect to the number of state variables. To reduce the number of state variables, we perform live variable analysis for the program up until the attack point - that is, live variable analysis is done assuming the program terminates after the attack point.
If the attack point is within one or more loops, then the assumed termination point would be immediately after all the loops are closed.
For our use case, we can treat the entire code till the assumed termination point as a single block. Any variables that are not live within this block are ignored, and only the remaining live variables are used as state variables for the signature extraction step.
Identifying live variables boils down to solving the dataflow equations:
$$ in[n] = use[n] \quad \bigcup  \quad (out[n] - def[n]) $$
$$ out[n] = \bigcup_{S \: \epsilon \: succ[n]}  in[s] $$

where $in$ and $out$ represent the live variables entering and leaving the edge in the control flow graph. $def$ and $use$ represent the variables defined and referenced in an expression represented by a node in the CFG.
\section{Signature Extraction} \label{sec:extraction}
We now describe a procedure to generate a network signature
of length $n$ (length of the signature is the number of packets
received to trigger attack) for a boolean malware program. 
We use the SMT solver Z3~\cite{z3}
to find valid network signatures. In this section, we describe
the constraints generated for the SMT solver from the program
such that the model generated by the solver can be used to 
extract signatures.

\subsection{Preliminaries}
For a malware program $M$, let $S$ denote
the vector of boolean state variables in the program,
and $P$ denote the vector of boolean packet variables
in the program\footnote{
	For example, a 60 bit packet is a vector 
	of 60 boolean variables.}. Let the
initial values of the state variables be $S_{0}$.

As described in the earlier section, the malware
packet processing code is a set of \texttt{if} 
statements whose predicates are over the set of state
and packet variables. Let the set of \texttt{if} 
statements be denoted by $I$, and let us denote each \texttt{if} 
statement by $(\gamma, \delta) \in I$, where $\gamma(S,P)$
denotes the predicate on $S$ and $P$, and $\delta(S,P)$ denotes
the action on $S$ if $\gamma(S,P)$ is satisfied. Therefore,
the semantics of the \texttt{if} statement is as follows:
\begin{equation}
\gamma(S,P) \Rightarrow S' == \delta(S,P)
\end{equation}
where $S'$ denotes the new state after the \texttt{if} 
block is executed.

Formally, given $M$, we need to find a sequence of packets $p_0$,
$p_1, \ldots p_{n-1}$ such that $M$ starting from initial
state reaches the attack point when it receives the packets in
that particular order. This sequence is packets is defined as a 
valid network signature of the malware program.

\subsection{Transition Model}
We denote the state
vector before receiving the $i^{th}$ packet as $S_i$.
When packet $i$ of the signature is received, 
mlaware $M$, based on $S_i$ and $P_{i}$, will execute 
a \texttt{if} $(\gamma_j, \delta_j)$ block.
For a given action block comprising  
assignment statements, the logical formula corresponding to
the assignments is: 
$\delta(S,P) = s_1' \Leftrightarrow \delta_1(S,P) \wedge s_2' \Leftrightarrow  \delta_2(S,P) \Leftrightarrow  \wedge ~s_{|S|}' = \delta_{|S|}(S,P)$, 
where $\delta_i$ corresponds to the right hand side
expression for the assignment statement $s_i = exp$.
If the action has no assignment for a state 
variable $s_i$, $\delta_i = s_i$ (the value must 
remain unchanged, so we need to add this clause).

We can model the malware state transition 
by the following constraints: 
\begin{equation}
\forall i. 0 \leq i \leq n-2.~\forall (\gamma, \delta) \in I. ~\gamma(S_i, P_{i}) \Rightarrow S_{i+1} \Leftrightarrow \delta(S_i, P_{i})
\end{equation}

To find minimal signatures, every packet received by the malware 
must trigger a state change, which translated to the constraint
that atleast one of the if predicates must be satisfied: 
\begin{equation}
\forall i. 0 \leq i \leq n-2. \bigvee_{(\gamma, \delta) \in I} \gamma(S_i, P_{i})
\end{equation}

Let the attack point block be $(\gamma_M, \delta_M)$. 
Therefore, a network signature of length $n$ means
that $P_{n-1}$ triggers the attack. The state of the program
is $S_{n-1}$. The following constraint must hold:
\begin{equation}
\gamma_M(S_{n-1},P_{n-1})
\end{equation}
The SMT solver will return a model for packets $P_1,
\ldots P_n$ such that the program after these receiving
these packets reaches the attack point. Therefore, 
we run this procedure for varying lengths $n$ to find
network signatures of different lengths.

Also, this procedure only returns a particular signature
of length $n$. We can generalize it by analyzing the 
transitions of $M$ to generalize the signature. For
example, if the $i^{th}$ packet only matches certain 
predicates in the transition, we can generalize the packet on
these bits (the rest of the bits are don't care). To 
generate the different singatures of length $n$, 
we can add blocking clauses for the signature we obtained 
to find a signature different from the current one.

\subsection{Example}
Let us consider the following program receiving two-bit packets:
\begin{lstlisting}	
s1 = 0;
s2 = 0;
while(True) {
packet = socket.receive();
if (s1 == 0 and packet[0] == 0) {
s1 = 1;
}
if (s1 == 1 and s2 == 0 and packet[1] == 1) {
s2 = (not s2) and packet[1];
}
if (s1 == 1 and s2 == 1 and packet[0] == 1 and packet[1] == 1) {
attack()  
}	
}
\end{lstlisting}	
We denote the state variables by $s_1, s_2$ and packet 
variables by $p_1, p_2$.
For the first \texttt{if} block, 
$\gamma: \neg s_1 \wedge \neg p_1$, and $\delta: s_1' \wedge s_2' \Leftrightarrow s_2$ ($s_1$ 
is set by the action). Similarily,
for the second block: 
$\gamma: s_1 \wedge \neg s_2 \wedge p_2$, and $\delta: 
s_1'\Leftrightarrow s_1 \wedge s_2' \Leftrightarrow \neg s_2 \wedge p_1$.
For this program, initial state $S_0 = (0, 0)$.
The attack point predicate $\gamma_M$ is as follows:
\begin{equation}
s_1 \wedge s_2 \wedge p_1 \wedge p_1
\end{equation}
Therefore, if we add the constraints as described in the 
earlier section, for $n=3$, we would obtain a
valid network signature.

\section{Evaluation}
We present some preliminary evaluation of signature extraction using
a synthetic malware example. We use the Z3~\cite{z3} SMT solver 
for solving the constraints generated for extracting signatures. 
We consider a malware program which processes 128 bit packets (i.e.,
128 packet boolean variables). The malware program contains $n$ state
variables which are initialized to false. For every packet satisfying 
a certain condition, one of the state variables is set to true, and the 
malware program launches an attack when all state variables are 
set to true. Therefore, the signature length is $n$. We show the 
time to solve constraints to extract the signature with varying 
signature lengths in \Cref{fig:eval}.
\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{plot.png}
	\caption{Time taken to extract a signature using the Z3 solver
		with varying signature lengths.}
	\label{fig:eval}
\end{figure}

\section{Future Work}
One of our simplifying assumptions is that we know the attack 
point in the malware. However, with large codebases, this may be 
difficult to identify manually. Therefore, an important extension is 
Identifying attack points automatically. Some initial approaches can
be identifying statements that make accesses to restricted files or 
running restricted commands (e.g., sudo). To detect propagation
in worms, 
we can find code which causes the malware to send itself on the 
network.

Another simplifying assumption is our malware program model 
as a set of if blocks in a while loop. We found real malware programs with
large codebases with mutliple files, thus packet processing can be 
across several procedures, requiring inter-procedural analyses.
Also, it may be difficult to find malware code in the wild, only binaries,
so extraction of signatures for a binary would be useful.

Another feature of malware code can be loops whose number 
of iterations depend on the bits of the packet. Thus, we cannot 
unroll the loop trivially, and require to find invariants for the loop 
predicated on packet characteristics.

Finally, the signature extraction can be used to verify  
implementations of receivers such that they adhere to 
a protocol specification.

\bibliographystyle{abbrv}
\bibliography{references}

\end{document}